Cloudbreak: A MapReduce Algorithm for Detecting Genomic Structural Variation

### 2013-03-18 manuscript/manuscript.tex
updated to use extract-in-hadoop calls

\begin{table}[t]
\begin{center}
\begin{tabular}{rrrrrr}
  \hline
 & 40-100bp  & 101-250bp  & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
 Total Number & 224 &  84 & 82 &  31 & 26\\ 
  \hline
  Cloudbreak  &   47 (   7)  &   50 (   2) &  \textbf{  55} (\textbf{   4}) & \textbf{  12} (\textbf{   4}) & \textbf{  15} (\textbf{0}) \\ 
  Breakdancer &   52 (  10)  &   49 (   2) &    49 (0) &    7 (0) &   14 (\textbf{0}) \\ 
  GASVPro     &   31 (   4)  &   25 (0) &    23 (0) &    2 (0) &    6 (\textbf{0}) \\ 
  DELLY       &   22 (   2)  & \textbf{  56} (\textbf{   3}) &    40 (0) &    8 (0) &   12 (\textbf{0}) \\ 
  Pindel      & \textbf{  60} (\textbf{  35})  &   16 (0) &    41 (   2) &    1 (0) &   12 (\textbf{0})\\ 
   \hline
\end{tabular}
\end{center}
\caption{The number of simulated deletions in the 30X diploid chromosome 2 with Venter indels found by each tool at a 10\% FDR, as well as the number of those deletions that were discovered exclusively by each tool (in parentheses). The total number of deletions in each size class in the true set of deletions is shown in the second row of the header.}
\label{chr2DeletionPredsFDR10}
\end{table}

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{rrrrrr}
  \hline
 & 40-100bp  & 101-250bp  & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
 Total Number & 224 &  84 & 82 &  31 & 26\\ 
  \hline
  Cloudbreak  & \textbf{ 155} (\textbf{  11})  &   57 (\textbf{   3}) &    65 (   2) &   15 (0) &   15 (0) \\ 
  Breakdancer &   89 (0)  &   54 (0) &    53 (0) &    8 (0) &   15 (0) \\ 
  GASVPro     &   83 (   2)  &   32 (0) &    55 (0) &    8 (0) &   15 (0) \\ 
  DELLY       &  114 (   7)  & \textbf{  68} (   2) &  \textbf{  66} (   1) &    9 (0) &   17 (   1) \\ 
  Pindel      &   96 (  10)  &   24 (0) &    48 (0) &    5 (0) &   15 (0)\\ 
  MoDIL      &  124 (   6)  &   66 (\textbf{   3}) &  \textbf{  66} (\textbf{   6}) & \textbf{  17} (\textbf{   4}) & \textbf{  23} (\textbf{   8})\\ 
   \hline
\end{tabular}
\end{center}
\caption{The number of simulated deletions in the 30X diploid chromosome 2 with Venter indels found by each tool at maximum sensitivity, as well as the number of those deletions that were discovered exclusively by each tool (in parentheses). The total number of deletions in each size class in the true set of deletions is shown in the second row of the header. TODO: Add precision and recall!!!!!}
\label{chr2DeletionPredsMaxSensitivity}
\end{table}

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{rrrrrr}
  \hline
 & 40-100bp  & 101-250bp  & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
 Total Number & 199 &  83 & 79 &  21 & 21\\ 
  \hline
  Cloudbreak  & \textbf{  65} (\textbf{  28})  &    9 (   5) &     3 (   2) &    1 (   1) & 0 (0) \\ 
  Breakdancer &   23 (   8)  & \textbf{  14} (\textbf{  10}) &     2 (   1) & 0 (0) & 0 (0) \\ 
  Pindel      &   52 (  26)  &    5 (   4) &  \textbf{  10} (\textbf{  10}) & \textbf{   3} (\textbf{   3}) & \textbf{   9} (\textbf{   9})\\ 
   \hline
\end{tabular}
\end{center}
\caption{The number of simulated insertions in the 30X diploid chromosome 2 with Venter indels found by each tool at maximum sensitivity, as well as the number of those insertions that were discovered exclusively by each tool (in parentheses). The total number of insertions in each size class in the true set of insertions is shown in the second row of the header. TODO: Add modil!!!! TODO: Add precision and recall!!!}
\label{chr2InsertionPreds}
\end{table}

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{rrr|rrrrr}
  \hline
 & Prec. & Recall & 40-100bp & 101-250bp & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
Total Number & & & 7,466 & 235 & 218 & 110 & 375 \\
  \hline
Cloudbreak & 0.0978 & \textbf{0.115} & \textbf{ 423} (\textbf{ 179})  &  128 (\textbf{   9}) &   158 (\textbf{   8}) &   70 (\textbf{   3}) &  186 (\textbf{  12}) \\ 
Breakdancer & 0.122 & 0.112 &  261 (  41)  & \textbf{ 132} (   8) &  \textbf{ 167} (   1) & \textbf{  92} (0) & \textbf{ 288} (  10) \\  
  GASVPro & 0.134 & 0.0401 &  104 (  17)  &   37 (   2) &    77 (0) &   26 (0) &   93 (0) \\ 
  DELLY & 0.0824 & 0.091 &  143 (   9)  &  125 (   7) &   158 (   1) &   83 (   1) &  256 (   3) \\ 
  Pindel & \textbf{0.16} & 0.0685 &  149 (  12)  &   57 (0) &   140 (   1) &   58 (0) &  172 (   2) \\ 
   \hline
\end{tabular}
\end{center}
\caption{The precision and recall with respect to the gold standard set of each tool on the NA18507 data, as well as the number of deletions in each size class found. The same cutoffs were used as for the simulated data as in Table \ref{chr2DeletionPredsFDR10}, adjusted for the difference in coverage from 30X to 37X.}
\label{NA18507DeletionPreds}
\end{table}

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{rrr|rrrrr}
  \hline
 & Prec. & Recall & 40-100bp & 101-250bp & 251-500bp & 501-1000bp \\ 
Total Number & & & 536 & 114 & 45 & 1 \\
  \hline
Cloudbreak & 0.0365 & 0.198 &  127 (  28)  &   11 (   3) &  0 (0) & 0 (0) \\ 
Breakdancer & 0.0104 & 0.181 &   97 (  36)  & \textbf{  27} (\textbf{  17}) &     2 (   2) & 0 (0) \\  
  Pindel & \textbf{0.0387} & \textbf{0.239} & \textbf{ 144} (\textbf{  76})  &   14 (  12) &  \textbf{   7} (\textbf{   7}) & \textbf{   1} (\textbf{   1})  \\ 
   \hline
\end{tabular}
\end{center}
\caption{The precision and recall with respect to the gold standard set of insertions for each tool on the NA18507 data, as well as the number of insertions in each size class found, at maximum sensitivity.}
\label{NA18507InsertionPreds}
\end{table}

@@

\begin{table}[b]
\begin{center}
\begin{tabular}{rrr|rr}
 & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\hline
 &  Non-repeat & Repeat  &  Non-repeat & Repeat \\ 
 Total Number & 120 & 327 & 553 & 7851 \\ 
  \hline
  Cloudbreak  &   28 (   4) & \textbf{ 151} (  13) & \textbf{ 204} (\textbf{  46}) & \textbf{ 761} (\textbf{ 165}) \\ 
  Breakdancer & \textbf{  29} (   5) &  142 (   7) &  186 (  21) &  754 (  39) \\
  GASVPro     &   15 (   2) &   72 (   2) &   71 (   6) &  266 (  13) \\
  DELLY       &   21 (   2) &  117 (   3) &  147 (  11) &  618 (  10) \\
  Pindel      &   18 (\textbf{   9}) &  112 (\textbf{  28}) &  103 (   4) &  473 (  11) \\ 
   \hline
\end{tabular}
\end{center}
\caption{Detected deletions on the simulated and NA18507 data sets identified by each tool, broken down by whether the deletion overlaps with a RepeatMasker-annotated element.}
\label{deletionRepmaskpreds}
\end{table}

\begin{table}[b]
\begin{center}
\begin{tabular}{rrr|rr}
 & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\hline
 &  Non-repeat & Repeat  &  Non-repeat & Repeat \\ 
 Total Number & 133 & 270 & 341 & 355 \\ 
  \hline
  Cloudbreak  &   19 (   5) & \textbf{  59} (  31) &   73 (  12) &   65 (  19) \\ 
  Breakdancer &   17 (   8) &   22 (  11) &   82 (  37) &   44 (  18) \\
  Pindel      & \textbf{  25} (\textbf{  19}) &   54 (\textbf{  33}) & \textbf{  84} (\textbf{  48}) & \textbf{  82} (\textbf{  48}) \\ 
   \hline
\end{tabular}
\end{center}
\caption{Detected insertions in the simulated and NA18507 data sets identified by each tool, broken down by whether the insertion occurs in a RepeatMasker-annotated element.}
\label{insertionRepmaskpreds}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{r|r|rr|rr|}
\multicolumn{2}{c}{}  & \multicolumn{4}{c}{Actual Genotypes} \\
\multicolumn{2}{c}{}  & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\cline{3-6}
\multicolumn{2}{c|}{} &  Homozygous & Heterozygous & Homozygous & Heterozygous \\ 
\cline{2-6}
\multirow{2}{*}{\shortstack{Predicted \\ Genotypes}} & Homozygous & 88 & 3 &  70 & 11 \\
 & Heterozygous & 18 & 70 &  4 & 209 \\
\cline{2-6}
\end{tabular}
\end{center}
\caption{Confusion matrices for the predicted genotype of deletions found by Cloudbreak on both the simulated and NA18507 data sets.}
\label{deletionGenotypeaccuracy}
\end{table}

>>>>

\begin{table}[t]
\begin{center}
\begin{tabular}{rrrrrr}
  \hline
 & 40-100bp  & 101-250bp  & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
 Total Number & 224 &  84 & 82 &  31 & 26\\ 
  \hline
  Cloudbreak  &   50 (   8)  &   50 (\textbf{   3}) &  \textbf{  55} (\textbf{   4}) & \textbf{  12} (\textbf{   4}) & \textbf{  15} (\textbf{0}) \\ 
  Breakdancer &   52 (   9)  &   49 (   2) &    49 (0) &    7 (0) &   14 (\textbf{0}) \\ 
  GASVPro     &   31 (   4)  &   25 (0) &    23 (0) &    2 (0) &    6 (\textbf{0}) \\ 
  DELLY       &   22 (   2)  & \textbf{  56} (\textbf{   3}) &    40 (0) &    8 (0) &   12 (\textbf{0}) \\ 
  Pindel      & \textbf{  60} (\textbf{  34})  &   16 (0) &    41 (   2) &    1 (0) &   12 (\textbf{0})\\ 
   \hline
\end{tabular}
\end{center}
\caption{The number of simulated deletions in the 30X diploid chromosome 2 with Venter indels found by each tool at a 10\% FDR, as well as the number of those deletions that were discovered exclusively by each tool (in parentheses). The total number of deletions in each size class in the true set of deletions is shown in the second row of the header.}
\label{chr2DeletionPredsFDR10}
\end{table}

@@

\begin{table}[b]
\begin{center}
\begin{tabular}{rrr|rrrrr}
  \hline
                        & Prec. & Recall & 40-100bp  & 101-250bp  & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
 Total Number &          &           & 224 &  84 & 82 &  31 & 26\\ 
  \hline
  Cloudbreak    &  \textbf{0.497} & 0.642 & \textbf{ 140} (   5)  &   58 (\textbf{   4}) &    61 (   1) &   13 (0) &   15 (0) \\ 
  Breakdancer   &  0.356 & 0.49 &   89 (0)  &   54 (0) &    53 (0) &    8 (0) &   15 (0) \\ 
  GASVPro        & 0.146 & 0.432 &   83 (   2)  &   32 (0) &    55 (0) &    8 (0) &   15 (0) \\ 
  DELLY           & 0.457 & 0.613 &  114 (   8)  & \textbf{  68} (   2) &  \textbf{  66} (   1) &    9 (0) &   17 (   1) \\ 
  Pindel           & 0.462 & 0.421 &   96 (\textbf{  14})  &   24 (0) &    48 (0) &    5 (0) &   15 (0)\\ 
  MoDIL           & 0.132  & \textbf{0.66} &  123 (   6)  &   66 (   2) &  \textbf{  66} (\textbf{   8}) & \textbf{  17} (\textbf{   6}) & \textbf{  23} (\textbf{   8})\\ 
   \hline
\end{tabular}
\end{center}
\caption{The number of simulated deletions in the 30X diploid chromosome 2 with Venter indels found by each tool at maximum sensitivy, as well as the number of those deletions that were discovered exclusively by each tool (in parentheses). The total number of deletions in each size class in the true set of deletions is shown in the second row of the header.}
\label{chr2DeletionPredsMaxSensitivity}
\end{table}

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{rrr|rrrrr}
  \hline
                        & Prec. & Recall & 40-100bp  & 101-250bp  & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
 Total Number &          &           & 199 &  83 & 79 &  21 & 21\\ 
  \hline
  Cloudbreak   &\textbf{0.624} & 0.194  & \textbf{  65} (  21)  &    9 (   4) &     3 (   1) &    1 (0) & 0 (0) \\ 
  Breakdancer & 0.262 & 0.0968  &   23 (   8)  & \textbf{  14} (\textbf{  10}) &     2 (   1) & 0 (0) & 0 (0) \\ 
  Pindel          & 0.572 & \textbf{0.196} &   52 (\textbf{  25})  &    5 (   3) &  \textbf{  10} (\textbf{  10}) & \textbf{   3} (\textbf{   3}) & \textbf{   9} (\textbf{   9})\\ 
  MoDIL          & 0.186 & 0.0521 &   14 (   1)  &    4 (   1) &     1 (0) &    2 (   1) & 0 (0)\\ 
   \hline
\end{tabular}
\end{center}
\caption{The number of simulated insertions in the 30X diploid chromosome 2 with Venter indels found by each tool at maximum sensitivity, as well as the number of those insertions that were discovered exclusively by each tool (in parentheses). The total number of insertions in each size class in the true set of insertions is shown in the second row of the header.}
\label{chr2InsertionPreds}
\end{table}

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{rrr|rrrrr}
  \hline
 & Prec. & Recall & 40-100bp & 101-250bp & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
Total Number & & & 7,466 & 235 & 218 & 110 & 375 \\
  \hline
Cloudbreak & 0.101 & \textbf{0.131} & \textbf{ 446} (\textbf{ 186})  & \textbf{ 141} (\textbf{  14}) &  \textbf{ 173} (\textbf{  12}) & \textbf{  93} (\textbf{   4}) &  248 (\textbf{  14}) \\ 
Breakdancer & 0.122 & 0.112 &  261 (  35)  &  132 (   6) &   167 (   1) &   92 (0) & \textbf{ 288} (   6) \\  
  GASVPro & 0.134 & 0.0401 &  104 (  17)  &   37 (   2) &    77 (0) &   26 (0) &   93 (0) \\ 
  DELLY & 0.0824 & 0.091 &  143 (   8)  &  125 (   6) &   158 (   1) &   83 (   1) &  256 (   3) \\ 
  Pindel & \textbf{0.16} & 0.0685 &  149 (  12)  &   57 (0) &   140 (0) &   58 (0) &  172 (   1) \\ 
   \hline
\end{tabular}
\end{center}
\caption{The precision and recall with respect to the gold standard set of each tool on the NA18507 data, as well as the number of deletions in each size class found. The same cutoffs were used as for the simulated data as in Table \ref{chr2DeletionPredsFDR10}, adjusted for the difference in coverage from 30X to 37X.}
\label{NA18507DeletionPreds}
\end{table}

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{rrr|rrrrr}
  \hline
 & Prec. & Recall & 40-100bp & 101-250bp & 251-500bp & 501-1000bp \\ 
Total Number & & & 536 & 114 & 45 & 1 \\
  \hline
Cloudbreak & \textbf{0.0596} & 0.197 &  124 (  28)  &   13 (   3) &  0 (0) & 0 (0) \\ 
Breakdancer & 0.0281 & 0.181 &   97 (  37)  & \textbf{  27} (\textbf{  15}) &     2 (   2) & 0 (0) \\  
  Pindel & 0.0387 & \textbf{0.239} & \textbf{ 144} (\textbf{  79})  &   14 (  12) &  \textbf{   7} (\textbf{   7}) & \textbf{   1} (\textbf{   1})  \\ 
   \hline
\end{tabular}
\end{center}
\caption{The precision and recall with respect to the gold standard set of insertions for each tool on the NA18507 data, as well as the number of insertions in each size class found, at maximum sensitivity.}
\label{NA18507InsertionPreds}
\end{table}

@@

\begin{table}[b]
\begin{center}
\begin{tabular}{rrr|rr}
 & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\hline
 &  Non-repeat & Repeat  &  Non-repeat & Repeat \\ 
 Total Number & 120 & 327 & 553 & 7851 \\ 
  \hline
  Cloudbreak  &   26 (   3) & \textbf{ 156} (  16) & \textbf{ 238} (\textbf{  61}) & \textbf{ 863} (\textbf{ 169}) \\ 
  Breakdancer & \textbf{  29} (   5) &  142 (   6) &  186 (  18) &  754 (  30) \\
  GASVPro     &   15 (   2) &   72 (   2) &   71 (   6) &  266 (  13) \\
  DELLY       &   21 (   2) &  117 (   3) &  147 (  10) &  618 (   9) \\
  Pindel      &   18 (\textbf{   9}) &  112 (\textbf{  27}) &  103 (   3) &  473 (  10) \\ 
   \hline
\end{tabular}
\end{center}
\caption{Detected deletions on the simulated and NA18507 data sets identified by each tool, broken down by whether the deletion overlaps with a RepeatMasker-annotated element.}
\label{deletionRepmaskpreds}
\end{table}

\begin{table}[b]
\begin{center}
\begin{tabular}{rrr|rr}
 & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\hline
 &  Non-repeat & Repeat  &  Non-repeat & Repeat \\ 
 Total Number & 133 & 270 & 341 & 355 \\ 
  \hline
  Cloudbreak  &   19 (   3) & \textbf{  59} (  23) &   75 (  13) &   62 (  18) \\ 
  Breakdancer &   17 (   8) &   22 (  11) &   82 (  36) &   44 (  18) \\
  Pindel      & \textbf{  25} (\textbf{  18}) &   54 (\textbf{  32}) & \textbf{  84} (\textbf{  48}) & \textbf{  82} (\textbf{  51}) \\ 
  MoDIL      &    5 (0) &   16 (   3) & NA & NA \\ 
   \hline
\end{tabular}
\end{center}
\caption{Detected insertions in the simulated and NA18507 data sets identified by each tool, broken down by whether the insertion occurs in a RepeatMasker-annotated element.}
\label{insertionRepmaskpreds}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{r|r|rr|rr|}
\multicolumn{2}{c}{}  & \multicolumn{4}{c}{Actual Genotypes} \\
\multicolumn{2}{c}{}  & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\cline{3-6}
\multicolumn{2}{c|}{} &  Homozygous & Heterozygous & Homozygous & Heterozygous \\ 
\cline{2-6}
\multirow{2}{*}{\shortstack{Predicted \\ Genotypes}} & Homozygous & 92 & 3 &  74 & 12 \\
 & Heterozygous & 15 & 71 &  4 & 286 \\
\cline{2-6}
\end{tabular}
\end{center}
\caption{Confusion matrices for the predicted genotype of deletions found by Cloudbreak on both the simulated and NA18507 data sets.}
\label{deletionGenotypeaccuracy}
\end{table}
